### 2022-09-29 ms/si.Qmd

---
title: "canaper: Categorical analysis of neo- and paleo-endemism in R"
subtitle: "Supporting information"
bibliography: references.yaml
csl: methods-in-ecology-and-evolution.csl
highlight-style: github
format:
  pdf:
    include-in-header:
      # keep figures from floating away
      text: |
        \usepackage{float}
        \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(targets)
library(tidyverse)
library(patchwork)
library(canaper)
```

```{r}
#| label: acacia-cpr
#| fig.height: 8
#| echo: false

acacia_canape <-
  tar_read(
    acacia_canape_cpr,
    store = here::here("_targets")
  ) |>
  separate(site, c("long", "lat"), sep = ":") |>
  mutate(across(c(long, lat), parse_number))

a <- ggplot(acacia_canape, aes(x = long, y = lat, fill = endem_type)) +
  geom_tile() +
  scale_fill_manual(values = cpr_endem_cols_4) +
  guides(
    fill = guide_legend(title.position = "top", label.position = "bottom")
  ) +
  coord_fixed() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = "white"),
    panel.grid.major = element_line(color = "grey60"),
    panel.grid.minor = element_blank()
  )

b <- ggplot(
  acacia_canape,
  aes(x = pe_alt_obs, y = pe_obs, color = endem_type)
) +
  geom_abline(slope = 1, color = "darkgrey") +
  geom_point() +
  scale_color_manual(values = cpr_endem_cols_4) +
  labs(
    x = "Phylogenetic endemism on comparison tree",
    y = "Phylogenetic endemism on actual tree"
  ) +
  theme_bw() +
  theme(legend.position = "none")

a + b + plot_layout(ncol = 1) + plot_annotation(tag_levels = "a")
```

Fig. S1: Categorical analysis of neo- and paleo-endemism (CANAPE) of Australian *Acacia*, plotted with color vision deficiency (CVD)-safe colors.
(a) Map of Australia showing grid-cells (communities) colored by endemism type.
(b) Scatterplot of comparing phylogenetic endemism (PE) of each community as measured on the original tree vs. a comparison tree with all non-zero branch lengths set to equal length, colored according to endemism type.